---
title: "Patient Profile"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = F, message = F, warning = F, fig.width = 12, fig.align="center")
```

# Using the patient profile

```{r libs}
library(haven)
library(data.table)
library(dplyr)
library(htmltools)
library(ggplot2)
```

```{r data}
AEf <- system.file("data/ADAE.sas7bdat", package = "pprofile")
CMf <- system.file("data/ADCM.sas7bdat", package = "pprofile")
SLf <- system.file("data/ADSL.sas7bdat", package = "pprofile")
LBf <- system.file("data/ADLB.sas7bdat", package = "pprofile")
EGf <- system.file("data/ADEG.sas7bdat", package = "pprofile")
EXf <- system.file("data/ADEX.sas7bdat", package = "pprofile")
PKf <- system.file("data/ADPC.sas7bdat", package = "pprofile")

AE <- read_sas(AEf)
CM <- read_sas(CMf)
SL <- read_sas(SLf)
LB <- read_sas(LBf)
EG <- read_sas(EGf)
EX <- read_sas(EXf)
PK <- read_sas(PKf)
```

# UI & Submodules

The patient profile only requires subject level data (ADSL). But new visualization
(submodules) become accessible when additional datasets are loaded.

Here is a description of the UI and the currently available submodules, along
with their data requirements.

## Patient summary
The landing page of the patient profile contains the patient selector, a quick
summary of the demographics & treatment information for the subject and a 
tabular overview of the available data.


## Adverse Events & Concomitant medication
If adverse events (ADAE) and concomitatn medications (ADCM) are available, the AE + CM
visualization is enabled. The plot displays the time ordered concomitant medications,
administration dates and duration, along with Adverse Event information and duration
also ordered by time.

```{r aecm, fig.height=9}
uid = "CXXX999Y9999_1000003"
aecm_list <- pprofile:::get_aecm_data(uid = uid, ADSL = SL, ADAE = AE, ADCM = CM)
pprofile:::plot_aecm(aecm_list = aecm_list, aecolour = "action")
```

Concomitant medications are drawn in purple. AEs can be colored by severity, grade, or 
action taken using the dropdown menu:

- Severity: colored by AE, or serious adverse event
- Grade: colored by whether the event is of Grade 1, 2 or 3
- Action taken: colored as to whether the dose remained unchanged, or whether there was an interruption in the dose 

The treatment period is highlighted with a green background.

This plot is particularly useful to investigate whether an observed AE is due to the 
drug under study or due to other medication.

## Exposure & Lab
When Lab values (ADLB), Exposure (ADEX) and PK data (ADPC) are available,
the “Exposure + Labs” tab is enabled. It provide the key liver toxicity markers
and dose administration record and calculated exposure information.  The lab
values are normalized to the upper limits of normality such that they may all
be displayed on the same graph. The lower graph shows each dosing of the
active compound and key PK exposure information.

```{r exlb, fig.height=7}
paramcd <- c('ALP', 'ALT', 'AST', 'BILI')
pprofile:::plot_EXLB(uid = uid, ADLB = LB, ADSL = SL, ADEX = EX, ADPC = PK, paramcd = paramcd)
```

## Lab Visits
If Lab values are given (ADLB), the module displays an overview of the parameters
that were collected for the participant and will highlight values that are
above or below the normal range.

```{r labo, fig.height=7}
labvis <- pprofile:::get_lab_values(uid = uid, ADLB = LB)
pprofile:::plot_lab_visits(labvis)
```

The lab overview can be used to quickly scan all available lab data for anomalies.
Once a parameter or timepoint of interest has been identified, it can be 
further investigated using the Lab Parameters tab.

## Lab Parameters
Always loaded along the Lab Visits, the Lab Parameters plot shows lab values over
time for the selected parameters. Upper Normal Limit (UNL) and Lower Normal Limit
(LNL) are displayed in red dashed lines.


```{r labp_ly, fig.height=7}
ply <- pprofile:::plotly_lab_param(uid, c("ALT", "AST", "ALP", "BILI"), ADLB = LB)
tagList(list(ply))
```

Parameters can be added or removed using a dropdown.



## Exposure & ECG
If Electrocardiogram data (ADEG), Exposure (ADEX) and PK (ADPC) are passed,
the Exposure & ECG tab will display the relationship among expose, ECG and 
dosing information.

```{r}
pprofile:::plot_EXEG(uid, param = "HRMEAN", ADSL = SL, ADEX = EX, ADEG = EG, ADPC = PK)
```

The ECG parameters to display are selected from a subject specific dropdown of
all available parameters.

## ECG
With Electrocardiogram data (ADEG) alone, the ECG submodule becomes available, 
without exposure information.

```{r ecg}
pprofile:::plot_ECG(uid, param = "HRMEAN", ADEG = EG)[["plot"]]
```

## Pharmacokinetics
If PK data (ADPC) is available, the PK module is enabled, providing the dosing
concentration over time, colored by analysis visit.

```{r pk}
pprofile:::plot_PK(uid, param = "Marker1 plasma concentrations (ng/ml)",
                   ADPC = PK)[["plot"]]
```

PK parameter to display can be selected from a dropdown.



# Implementation of participant profile
If you like what you see and would like to build an app for your data
that includes the patient profile or add the module to an existing app, 
follow the documentation in [patprofile for developers](./pprofile_dev.html) tab.

You can also contact the [SCC](http://go/scc) for information about the module or
collaboration for app development.
