---
title: "Patient profile for developers"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: cerulean
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE, eval = FALSE, message = FALSE)
```

# Adding the patient profile to an app


The patient profile has been packaged as a [shiny module](https://shiny.rstudio.com/articles/modules.html) to facilitate 
deployment. It can be added to any app that uses standard ADaM datasets with 
only a couple of lines. 

#### Server

Server side, ADaM datasets are passed in the module call.
```{r, eval = F}
callModule(patientProfile_mod, id = "pp_module1",
           ADSL = myADSL, ADAE = myADAE, ADCM = myADCM, ADLB = myADLB,
           ADEG = myADEG, ADEX = myADEX, ADPC = myADPC)
```
Aside from `ADSL`, all other datasets can be left NULL. Missing datasets
simply disable some of the functionalities of the module. Moreover, 
whenever informative columns are missing, the submodules will attempt to
find alternative sources within the given data.


#### UI

UI side, `patientProfile_modUI` can be used to 
```{r, eval = F}
patientProfile_modUI(id = "pp_module1")
```


# Submodules
While the patient profile is meant to be used as a single top level module that
will adapt to the data available, submodules are exported by the package and
can be used individually. They are currently impplemented as `tabPanel`, so 
their use is restricted to compatible containers (`navbarPage`/`avaNavbarPage`,
`tabsetPanel`/`avaTabPanel`, ...).


For example, to include the Exposure & Lab values module alone
```{r, eval = F}
# Server
callModule(module = exlb_mod, id = "exlb_auto", uid = uid,
           ADSL = ADSL, ADLB = ADLB, ADEX = ADEX, ADPC = ADPC)
# UI (within a tabset or menubar)
exlb_UI(ns("exlb_auto"))
```

When used in this manner, it is the developer's responsibility to add the 
appropritate data checks to ensure correct behaviour.

Refer to the individual submodules documentation for details on the 
required inputs.

# Linking patient profile with other modules

A `uid` argument can be passed to the patient profile. By creating a new
input binding in the server code and linking it to a `Shiny.onInputChange`
event, we can set the participant profile to automatically 
select a participant of interest targeted from another plot.

```{r}
# Server code
uid <- reactiveVal()
observe({
  ns <- session$ns
  if(!is.null(input$patient_js)){
    print("Updating selected patient")
    uid(input$patient_selector)
  }
})
# Module call
callModule(patientProfile_mod, id = "pp_module1", uid = uid, ADSL = myADSL)
```


## Existing implementations
The [AVA safety explorer](http://go/scc/ase-pkg) is linked with the patient profile, such that when a
participant is clicked in the timelines or safety outlier explorer, the 
subject is pre-selected in the patient profile module.

Likewise, if the [eDISH](https://github.com/SafetyGraphics/hep-explorer) plot is
also part of the app, it will have the same interaction with the participant 
profile.

## Custom plots 
An example is available in the `pp-template` shiny app embeded within the package.

Using `plotly`, we can setup a plot with patient ids as key.
```{r, eval = F}
p <- plot_ly(data = data, x = ~ADT, y = ~AVAL,
             key = ~USUBJID,
             type = "scatter", mode = "markers",
             text = ~paste("uid:", USUBJID))
```

With the `onRender` function of
the `htmlwidgets` package, javascript code can be added to plot events. We can
use this to link a custom plot with the patient profile
```{r, eval = F}
# p is a plotly object
p <- p %>% onRender("
  function(el){
    el.on('plotly_click', function(d){
      selsub = d.points[0].data.key[d.points[0].pointNumber];
      console.log('Click: ', selsub);
      Shiny.onInputChange('pp_module1-patient_js', selsub);
    })
  }
")
```

The chunk of code assume that the module id is "pp_module1". If the module is
named differently, the render string can be built 

```{r, eval = F}
module_id <- "pp_module2"
render <- paste0("
  function(el){
    el.on('plotly_click', function(d){
      selsub = d.points[0].data.key[d.points[0].pointNumber];
      console.log('Click: ', selsub);
      Shiny.onInputChange('", module_id, "-patient_js', selsub);
    })
  }
")
```

If the plot function is part of its own module, then the namespace must be
passed instead of the module name.

```{r, eval = F}
target <- ns("patient_js")
render <- paste0("
  function(el){
    el.on('plotly_click', function(d){
      selsub = d.points[0].data.key[d.points[0].pointNumber];
      console.log('Click: ', selsub);
      Shiny.onInputChange('", target, "', selsub);
    })
  }
")
```


# Creating submodules for patprofile

All visualization tools were created as self contained submodules, making it
trivial to extend the functionalities of the patient profile by adding new
submodules.


Developers are encouraged to submit merge requests with their own
shiny modules or open issues to suggest new visualization on the
[package's GitLab page](https://gitlabce.statwb.eu.novartis.net/SAUTERE1/pprofile).
